Improving Day Ahead Electricity Load Forecasts with Google Trends

### 2014-12-04 Paper/DraftPaper.Rnw

\begin{abstract}
Modern short term load forecasting has grown in analytically complexity and sophistication.  Day ahead forecasts now commonly use neural nets, Monte Carlo simulations and a wealth of historical data.  What they have not done is fully captured the sentiment and intentions of the people using the electricity.  This paper introduces Google Trend data, a summary of Google searches, as a way of capturing this sentiment and refining forecasts.  We show with drop all forward cross validation that this amendment decreases forecast uncertainty by approximately 5\% when compared to a statistically adjusted forecast and by over 50\% when compared to raw forecasts.
\end{abstract}

@@

<<echo=FALSE,results=hide>>=
load("~/Research/Forecast-Peak-With-Google-Trends/DataManipulation/WTrends")

#Get rid of the zero forecast in hour two
WTrends$F02[WTrends$F02==0]<-NA

WTrends$Weather<-((925749*WTrends$DETrends)+(4395295*WTrends$KYTrends)+(5928814*WTrends$MDTrends)+(8899339*WTrends$NJTrends)+(11570808*WTrends$OHTrends)+(12773801*WTrends$PATrends)+(8260405*WTrends$VATrends)+(1854304*WTrends$WVTrends))/(925749+4395295+5928814+8899339+11570808+12773801+8260405+1854304)  
@

\begin{enumerate}
  \item Intro to short term load forecasting.
  \item Why crowd sourced, non technical,  information could be useful.
  \item Google trends is the summation of Google searches.
  \item Outline of paper
\end{enumerate}

@@

<<echo=false,results=hide>>=
library(nlme)
HourModelForecastCheck<-function(hour){
  Hour<-formatC(hour, width=2, flag="0")
  as.formula(paste("HE",Hour,"~ F",Hour,sep='' ))}

SAResults<-lapply(1:24, FUN = function(x)  gls(HourModelForecastCheck(x), data=WTrends,corAR1(form=~1),na.action=na.omit))

@

\begin{figure}
\begin{center}
\caption{Confidence Intervals for Intercept Statistically Adjusted Models (95\%)}
<<echo=false,fig=TRUE>>=
library(ggplot2)
LowInter<-unlist(lapply(SAResults, FUN= function(x) confint(x) [1,1]))
HighInter<-unlist(lapply(SAResults, FUN= function(x) confint(x) [1,2]))
InterceptResults<-data.frame(1:24,LowInter,HighInter)
names(InterceptResults)[1]<-"Hour"

ggplot(InterceptResults, aes(Hour))+geom_pointrange(aes(y=(LowInter+HighInter)/2, ymin=LowInter,ymax=HighInter))+geom_abline(intercept = 0)+ylab("Intercept Confidence Interval")
@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Confidence Intervals for Co-Movement Statistically Adjusted Models (95\%)}
<<echo=false,fig=TRUE>>=
library(ggplot2)
LowInter<-unlist(lapply(SAResults, FUN= function(x) confint(x) [2,1]))
HighInter<-unlist(lapply(SAResults, FUN= function(x) confint(x) [2,2]))
InterceptResults<-data.frame(1:24,LowInter,HighInter)
names(InterceptResults)[1]<-"Hour"

ggplot(InterceptResults, aes(Hour))+geom_pointrange(aes(y=(LowInter+HighInter)/2, ymin=LowInter,ymax=HighInter))+geom_abline(intercept = 1,slope=0)+ylab("Intercept Confidence Interval")
@
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\caption{State Weather Trends Indexes Over Time}
<<echo=false,fig=TRUE>>=
library(ggplot2)
library(reshape2)

JustWeather<-data.frame(WTrends$Date, WTrends$DETrends,WTrends$KYTrends, WTrends$MDTrends, WTrends$NJTrends, WTrends$OHTrends, WTrends$PATrends, WTrends$VATrends, WTrends$WVTrends, WTrends$Weather)

names(JustWeather)<-c("Date","DE","KY","MD","NJ","OH","PA","VA","WV","Weather")

MeltedTrends<-melt(JustWeather,variable.name="Index",id.var="Date",na.rm=TRUE)

ggplot(MeltedTrends, aes(Date, value,group= Index))+geom_line() +ylab("Index")+facet_grid(Index~.)

@
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\caption{Trends Indexes Over Time}
<<echo=false,fig=TRUE>>=
library(ggplot2)
library(reshape2)

JustTrends<-data.frame(WTrends$Date, WTrends$GasTrends,WTrends$MovieTrends, WTrends$NewsTrends, WTrends$RestaurantTrends, WTrends$TrafficTrends,WTrends$Weather)
names(JustTrends)<-c("Date","Gas","Movie","News","Restaurant","Traffic","Weather")

MeltedTrends<-melt(JustTrends,variable.name="Index",id.var="Date",na.rm=TRUE)

ggplot(MeltedTrends, aes(Date, value,group= Index))+geom_line() +ylab("Index")+facet_grid(Index~.)

@
\end{center}
\end{figure}

@@

<<echo=false,results=hide>>=
WTrendsLimited<-WTrends[!is.na(WTrends$Weather),]

HourModel<-function(hour){
  Hour<-formatC(hour, width=2, flag="0")
  as.formula(paste("HE",Hour,"~ F",Hour,"+Weather",sep='' ))}

# Gives percent reduction in sigma
CompareRaw<-function(x){
SA<-sd(WTrendsLimited[,x]-WTrendsLimited[,30])  
Google<-summary(lm(HourModel(x), data=WTrendsLimited))$sigma

(SA-Google)/(SA)
}

DirectCompare<-lapply(1:24, FUN = CompareRaw)
mean(unlist(lapply(1:24, FUN = CompareRaw)))

@

<<echo=false,results=hide>>=
HourModel<-function(hour){
  Hour<-formatC(hour, width=2, flag="0")
  as.formula(paste("HE",Hour,"~ F",Hour,"+Weather",sep='' ))}

GrossTrends<-lapply(1:24, FUN = function(x)  gls(HourModel(x), data=WTrends,corAR1(form=~1),na.action=na.omit))
@

\begin{figure}
\begin{center}
\caption{Confidence Intervals for ``Weather'' in Trends Models (95\%)}
<<echo=false,fig=TRUE>>=
library(ggplot2)

LowInter<-unlist(lapply(GrossTrends, FUN= function(x) confint(x) [3,1]))
HighInter<-unlist(lapply(GrossTrends, FUN= function(x) confint(x) [3,2]))
InterceptResults<-data.frame(1:24,LowInter,HighInter)
names(InterceptResults)[1]<-"Hour"

ggplot(InterceptResults, aes(Hour))+geom_pointrange(aes(y=(LowInter+HighInter)/2, ymin=LowInter,ymax=HighInter))+geom_abline(intercept = 1,slope=0)+ylab("Confidence Interval")
@
\end{center}
\end{figure}

<<echo=false,results=hide>>=

# Gives percent reduction in sigma
CompareSigma<-function(x){
SA<-summary(lm(HourModelForecastCheck(x), data=WTrendsLimited))$sigma  
Google<-summary(lm(HourModel(x), data=WTrendsLimited))$sigma

(SA-Google)/(SA)
}

ModelCompare<-lapply(1:24, FUN = CompareSigma)
mean(unlist(lapply(1:24, FUN = CompareSigma)))

@

@@

<<echo=false,results=hide>>=

OneForwardError<-function(formula, nData){
  model<-gls(formula, data=WTrendsLimited[1:nData,],corAR1(form=~1),na.action=na.omit)
#   modelPlus1<-lm(formula, data=WTrendsLimited[1:(nData+1),], y=TRUE)
#   predict(model,WTrendsLimited[1+nData,])-tail(modelPlus1$y,1)

tail(resid(model),1)

}

CVDropForward<-function(model){sd(unlist(lapply(20:(length(WTrendsLimited)-1), FUN= function(x) OneForwardError(model,x))))}

CompareSigmaCV<-function(x){
SA<-CVDropForward(HourModelForecastCheck(x))  
Google<-CVDropForward(HourModel(x))

(SA-Google)/(SA)
}

CVImprove<-lapply(1:24, FUN = CompareSigmaCV)
CVImprove
mean(unlist(CVImprove))
@

<<echo=FALSE,results=tex>>=
ForCompareTable<-data.frame(1:24,unlist(DirectCompare)*100,unlist(ModelCompare)*100,unlist(CVImprove)*100)
names(ForCompareTable)<-c("Hour","Direct","Statistically Adjusted (Raw)", "Statistically Adjusted (CV)")
library(stargazer)
stargazer(ForCompareTable,summary=FALSE, title="Improvement in Forecasts Relative to Gross,  Statistically Adjusted, Drop Forward CV (Percent)",rownames=FALSE)
@

@@

<<echo=false,results=tex>>=
News<-gls(HE19~F19+NewsTrends,data=WTrends,corAR1(form=~1),na.action=na.omit)
# Oh, that works too
Gas<-gls(HE19~F19+GasTrends,data=WTrends,corAR1(form=~1),na.action=na.omit)
# Gas does not
Traffic<-gls(HE19~F19+TrafficTrends,data=WTrends,corAR1(form=~1),na.action=na.omit)
# Traffic does not
Restaurant<-gls(HE19~F19+RestaurantTrends,data=WTrends,corAR1(form=~1),na.action=na.omit)
# Restaurant does not
Movie<-gls(HE19~F19+MovieTrends,data=WTrends,corAR1(form=~1),na.action=na.omit)
# Snort...Movie does.  Cool place to be.
library(stargazer)
stargazer(News,Gas,Traffic,Restaurant,Movie,title="Alternate Google Search Models for Hour 19",column.labels=c("News","Gas","Traffic","Restaurant","Movie"), dep.var.caption="Hour 19 Load",dep.var.labels.include = FALSE)
@

@@

<<echo=false,results=tex>>=
# 
# for(x in 1:12){
#   stargazer(GrossTrends[[x]], title=paste("Hour",x), dep.var.labels=paste("Hour",x), covariate.labels=c("Forecast","Weather","Constant") )
# }
@
\clearpage

<<echo=false, results=tex>>=
# for(x in 13:24){
#   stargazer(GrossTrends[[x]], title=paste("Hour",x), dep.var.labels=paste("Hour",x), covariate.labels=c("Forecast","Weather","Constant") )
# }
@

>>>>

\begin{abstract}
Modern short term load forecasting has grown in analytically complexity and sophistication.  Day ahead forecasts now commonly use neural nets, Monte Carlo simulations and a wealth of historical data.  What they have not done is fully captured the sentiment and intentions of the people using the electricity.  This paper introduces Google Trend data, a summary of Google searches, as a way of capturing this sentiment and refining forecasts.  We show with drop all forward cross validation that this amendment decreases forecast uncertainty when compared to a statistically adjusted forecast in the PJM control area.
\end{abstract}

@@

\begin{enumerate}
  \item Why and the research question why it is important, i.e., ancilary services cost.
  \item Intro to short term load forecasting.
  \item Why crowd sourced, non technical,  information could be useful.
  \item Other places where trends data has been useful.
  \item Google trends is the summation of Google searches.
  \item Outline of paper
\end{enumerate}

@@

@@

@@

@@

@@

@@
